---
title: "p8105_hw6_hc3654"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)
```


# Problem 1:

```{r}
cities_df <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv")
```

## Clean data

```{r}
cities_clean_df <- cities_df |> 
  mutate(city_state = paste(city,state,sep = ", "), 
         solve = case_when(
          disposition == "Closed without arrest" ~ 0, 
          disposition == "Open/No arrest" ~ 0,
          disposition == "Closed by arrest" ~ 1)) |> 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  mutate(victim_age = as.numeric(victim_age
  )) 
  
```

## Regression 

```{r}
fit_log = cities_clean_df |> 
  filter(city_state == "Baltimore, MD") |> 
  glm(solve ~ victim_race + victim_sex + victim_age, data = _, family = binomial()) |> 
  broom::tidy(conf.int = TRUE)

fit_log

```

```{r}
fit_log |> 
   filter(term == "victim_sexMale") |> 
  mutate(OR = exp(estimate), OR_lower = exp(conf.low), OR_higher = exp(conf.high)) |>
  select(term, OR, OR_lower, OR_higher) |> 
  knitr::kable(digits = 3)
```

## For all cities 

```{r}


odds_ratios_cities <- cities_clean_df |>
  group_by(city_state) |>
  nest() |> 
  mutate(
    model_results = purrr::map(
      .x = data,
      .f = ~ glm(solve ~ victim_race + victim_sex + victim_age, 
                 data = .x, 
                 family = binomial()) |>
             broom::tidy(conf.int = TRUE))) |> 
      unnest(model_results) |> 
filter(term == "victim_sexMale") |> 
  mutate(OR = exp(estimate), OR_lower = exp(conf.low), OR_higher = exp(conf.high)) |>
  select(term, OR, OR_lower, OR_higher) |> 
  arrange(OR)

odds_ratios_cities |> 
  select(-term)

odds_ratios_cities|> 
ggplot(aes(x = OR, y = fct_reorder(city_state, OR)))+
  geom_point()+
  geom_errorbarh(aes(xmin = OR_lower, xmax = OR_higher), 
                 height = 0.2, 
                 color = "darkgray")
```
### Comment:

The plot reveals that the adjusted Odds Ratio (OR) is less than 1 for the majority of cities, indicating a prevailing pattern where female victims' homicides have a higher solved advantage (odds) than those of male victims.

However, a crucial finding is that the confidence interval for the OR includes 1 for most cities. This means we lack sufficient statistical evidence to claim a significant gender difference in clearance rates in the majority of locations.

The results highlight two extremes: New York, NY, exhibits the strongest disparity (OR lowest), where the difference in solved odds between male and female victims is most pronounced and statistically significant. Conversely, Albuquerque, NM, holds the highest OR, suggesting male victims' cases are either equally or more likely to be solved there, it's not significant due to including the 1.0.

The clustering of most OR estimates between 0.5 and 1.0 confirms that while a bias exists, the magnitude of the disadvantage for male victims is generally moderate to significant across the cities, driving the overall conclusion that female victims' homicides are generally more likely to be solved across the dataset.


# Problem 2:

```{r}
library(p8105.datasets)
data("weather_df")
```


```{r}
weather_cleaned_df = weather_df |> 
  na.omit()

fit_t <- lm(tmax ~ tmin + prcp, data = weather_cleaned_df) 

fit_t |> 
  broom::tidy() |> 
  filter(term == "tmin"| term == "prcp") |> 
  select(term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(ratio = tmin/prcp) |> 
  select(ratio)


fit_t |> 
  broom::glance() |> 
  select(r.squared)

```


```{r}
set.seed(1)

fit_bootstrap = 
  weather_cleaned_df |> 
  bootstrap(n = 5000) |> 
  mutate(
    df = map(strap, as_tibble),
    fits = map(df, \(df) lm(tmax ~ tmin + prcp, data = df))
  ) |> 
  mutate(r_squared = map_dbl(fits, 
                        .f = ~ broom::glance(.x) |> pull(r.squared)
    )) |>
  mutate(
      beta_ratio = map_dbl(fits, 
                         .f = ~ broom::tidy(.x) |> 
                           filter(term %in% c("tmin", "prcp")) |> 
                           select(term, estimate) |> 
                           pivot_wider(names_from = term, values_from = estimate) |> 
                           mutate(ratio = tmin / prcp) |>
                           pull(ratio))) |> 
  select(.id, r_squared, beta_ratio)

```

### For R^2 

```{r}
ggplot(fit_bootstrap, aes(x = r_squared)) +
  geom_density(fill = "lightblue", alpha = 0.6) +
  labs(title = "Bootstrap Distribution of R-squared", x = "R^2 Estimate") +
  theme_minimal()


fit_bootstrap |>
  summarize(
    R2_lower = quantile(r_squared, 0.025),
    R2_upper = quantile(r_squared, 0.975)
  ) |> 
    knitr::kable(digits = 3)
```


### For Beta(1) / Beta(2)

```{r}
ggplot(fit_bootstrap, aes(x = beta_ratio)) +
  geom_density(fill = "lightblue", alpha = 0.6) +
  labs(title = "Bootstrap Distribution of R-squared", x = "R^2 Estimate") +
  theme_minimal()


fit_bootstrap |>
  summarize(
    Ratio_lower = quantile(beta_ratio, 0.025, na.rm = TRUE),
    Ratio_upper = quantile(beta_ratio, 0.975, na.rm = TRUE)
  ) |> 
    knitr::kable(digits = 3)
```


# Problem 3:

```{r}
baby_df_cleaned = read.csv("birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(babysex = as.factor(babysex), frace = as.factor(frace), malform = as.factor(malform), mrace = as.factor(mrace), parity = as.factor(parity), smoken = as.factor(smoken)) |> 
  drop_na()


my_mod <- lm(bwt ~ mheight + delwt + mrace, data = baby_df_cleaned) 


baby_df_cleaned|> 
  add_predictions(my_mod) |> 
  add_residuals(my_mod) |> 
  ggplot(aes(x = pred, y = resid))+
  geom_point(alpha = 0.5)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")+
  labs(title = "residuals vs. predictions",
       subtitle = "model: bwt ~ mheight + delwt + mrace ",
       x = "predictions",
       y = "residuals") +
  theme_bw()
```
```{r}
# Other two models
main_effs_mod = lm(bwt ~ blength + gaweeks, data = baby_df_cleaned)

three_ways_mod = lm(bwt ~ bhead + blength+ babysex, data = baby_df_cleaned)
```


```{r}
# Comparison 

cv_df = 
  modelr::crossv_mc(baby_df_cleaned, n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df_fits <- cv_df |>
  mutate(
    my_mod_fit = map(train, \(df) lm(my_mod, data = df)),
    main_effs_mod_fit = map(train, \(df) lm(main_effs_mod, data = df)),
    three_ways_mod_fit = map(train, \(df) lm(three_ways_mod, data = df))
  ) |> 
  mutate(
    rmse_my_mod = map2_dbl(my_mod_fit, test, modelr::rmse),
    rmse_main_effs = map2_dbl(main_effs_mod_fit, test, modelr::rmse),
    rmse_three_ways = map2_dbl(three_ways_mod_fit, test, modelr::rmse)
  )


cv_df_fits |>
  select(starts_with("rmse_")) |>
  pivot_longer(everything(), names_to = "model", values_to = "rmse") |>
  group_by(model) |>
  summarize(
    rmse = mean(rmse),
    .groups = "drop"
  ) 


cv_df_fits |> 
  select(rmse_main_effs, rmse_three_ways, rmse_my_mod) |> 
  pivot_longer(
    everything(),
    names_to = "mod",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x = mod, y = rmse, fill = mod))+
  geom_violin()
```


